---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
setwd("~/Documents/Coursera/5 - Reproducible research/Assignment 1/RepData_PeerAssessment1")
data_raw <- read.csv("activity.csv")
data <- data_raw[complete.cases(data_raw),]
```

## What is mean total number of steps taken per day?

```{r}
library(dplyr)

stepsbyday <- data %>% group_by(date) %>% summarize(steps= sum(steps))

hist(stepsbyday$steps,breaks=20, main= "Number of steps per day", xlab = "Steps")

abline(v=mean(stepsbyday$steps),col="red")
abline(v=median(stepsbyday$steps),col="blue")
legend("topright", c("Mean","Median"),lty=1, col=c("red","blue"))
```

The mean and the median are almost indistinguishable:
```{r}
mean(stepsbyday$steps)
median(stepsbyday$steps)
```

## What is the average daily activity pattern?

```{r}
stepsbyhour <- data %>% group_by(interval) %>% summarize (steps = mean(steps))
```

The maximum is reached for the interval:
```{r}
maxi <-stepsbyhour$interval[which.max(stepsbyhour$steps)]
print(maxi)

plot(stepsbyhour, type = "l")
abline(v=maxi,col="red")
```

## Imputing missing values
Number and percentage of missing values:
```{r}
sum(is.na(data_raw$steps))
sum(is.na(data_raw$steps)) / length(data_raw$steps) * 100
```

A look at the data shows that NA values come by entire days. We can then replace this missing days by an "average" day (computed previously in **stepsbyhour**).

```{r}
where <- which(is.na(data_raw$steps))

data_filled <- data_raw    # copy
for (i in where)
{
  missing_interval <- data_raw$interval[i]
  data_filled$steps[i] <-stepsbyhour$steps[ stepsbyhour$interval == missing_interval ]
}

```

Checking if it worked:
```{r}
head(data_raw)
head(data_filled)
```
and if we didn't mess everything up:
```{r}
data_raw[1000:1010,]
data_filled[1000:1010,]
```

Let's have a look at the new data:
```{r}
stepsbyday_filled <- data_filled %>% group_by(date) %>% summarize(steps= sum(steps))

hist(stepsbyday_filled$steps,breaks=20, main= "Number of steps per day", xlab = "Steps")

abline(v=mean(stepsbyday_filled$steps),col="red")
abline(v=median(stepsbyday_filled$steps),col="blue")
legend("topright", c("Mean","Median"),lty=1, col=c("red","blue"))

mean(stepsbyday_filled$steps)
median(stepsbyday_filled$steps)
```
As a reminder, dropping the NAs yielded:
```{r}
mean(stepsbyday$steps)
median(stepsbyday$steps)
```

The mean hasn't moved at all: we only completed the previous data with an average day, preserving the mean.
The median has slightly changed and now equals the mean.


## Are there differences in activity patterns between weekdays and weekends?

My R install is in French so weekends are *samedi* (Saturday) and *dimanche* (Sunday):
```{r}
data_filled$daytype <- weekdays(as.Date(data_filled$date)) %in% c("samedi","dimanche")
data_filled$daytype <- factor(data_filled$daytype, levels=c("TRUE","FALSE"),label=c("weekend","weekday"))

```


```{r}
stepsbyhour_filled <- data_filled %>% group_by(daytype,interval) %>% summarize (steps = mean(steps))

library(lattice)
xyplot(steps~interval| daytype, stepsbyhour_filled, type="l", layout=c(1,2))
```

It looks like he wakes up early during the weekdays to go and sit all day long: he has an office job.
